---
title: "215 Lab 2"
author: "Laura Wolf"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(entropy)

df <- as.data.frame(read_csv("W19/215/data/ex2_sample_data.csv"))

```
Mutual information: Look at set of variables and the shared information between them. Want to see where two variables are occurring together in space (where they share information). Look at two descriptor variables at a time. 
geology, radiation, and flow. 

```{r geol}

table(df[,3])
table(df[,6])
table(df[,3],df[,6])
mi.empirical(table(df[,6],df[,3])) #mutual information
entropy(df[,6]) #entropy
entropy(df[,3]) #entropy

```

Features of geology and vegetation co-occur 16% of the time

```{r veg_rad}

table(df[,4])
table(df[,6])
mi.empirical(table(df[,6],df[,4])) #mutual information
entropy(df[,6]) #entropy
entropy(df[,4]) #entropy

```

```{r veg_flow}

table(df[,5])
table(df[,6])
mi.empirical(table(df[,6],df[,5])) #mutual information
entropy(df[,6]) #entropy
entropy(df[,5]) #entropy

```

At the first level, geology is the strongest predictor of vegetation or has the most mutual information with vegetation

```{r geology_flow}

geol1_vegflow <- mi.empirical(table(df[df[,3]==1,6],df[df[,3]==1,5]))
geol1_vegflow #0.008374299
table_geoflow <- table(df[df[,3]==1,6],df[df[,3]==1,5])
table_geoflow # frequency of veg classes with geology type 1 in each flow type

geol2_vegflow <- mi.empirical(table(df[df[,3]==2,6],df[df[,3]==2,5]))
geol2_vegflow #0.008374299

geol5_vegflow <- mi.empirical(table(df[df[,3]==5,6],df[df[,3]==5,5]))
geol5_vegflow #0.005408553

geol6_vegflow <- mi.empirical(table(df[df[,3]==6,6],df[df[,3]==6,5]))
geol6_vegflow #0.009004135


```

```{r geology_rad}

#geology and radiation

geol1_vegrad <- mi.empirical(table(df[df[,3]==1,6],df[df[,3]==1,4]))
geol1_vegrad #0.06348332 This is the highest number.

table_georad <- table(df[df[,3]==1,6],df[df[,3]==1,4])
table_georad # frequency of veg classes with geology type 1 in each flow type

geol2_vegrad <- mi.empirical(table(df[df[,3]==2,6],df[df[,3]==2,4]))
geol2_vegrad #0.02894113

geol5_vegrad <- mi.empirical(table(df[df[,3]==5,6],df[df[,3]==5,4]))
geol5_vegrad #0.03346215

geol6_vegrad <- mi.empirical(table(df[df[,3]==6,6],df[df[,3]==6,4]))
geol6_vegrad #0.04961867

```

